\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-02-19 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
keep on updating

\title{ \blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}}

@@

\begin{abstract} 
\KCnote{Numerical modelling of multiphase flow in porous media has a wide range of applications in science and engineering. Viscous fingering, a fluid instability can be a major concern regarding enhanced oil-recovery methods (EOR). Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics and tackle the difficulties that commercial resevoir simulators may rise when using low order finite volume/differences methods based on structured grids. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations. We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.
In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion. Both the evolution of visualisation techniques through the extensive use of parallel computing as well as the advances in modelling and simulation of porous media flows may help to address the importance of this phenomenon which can be found in petroleum, chemical and environmental operations.
The aim of this paper is to extend and further apply the formulation (embedded in the Fluidity software framework) to multiphase flows in heterogeneous porous media combined with either uniform or random generated permeability distribution. So far, the results illustrate that the displacing fluid tends to flow easier towards the zones with higher values of permeability (spatial acceleration / deceleration of flow – fingers formation). It can also be seen that finger like structure will remain connected as it travels towards the outlet while it will grow faster in zones with higher permeability. This behaviour is triggered by the permeability contrast among layers.
As can be seen from Fig........ the higher the viscosity ratio the easier it is for the uniform front of the injected fluid to break into fingers that will decrease the recovery of hydrocarbon content. Also, the higher the viscosity ratio, more to the right the breakpoint is shifted, in the time axis. The increased viscosity ratio increases the computational cost and requires the decrease of the solution's time-step.
Both the geometrical domain and the development of the flow, as can be seen in Fig........ follow the linear tip-splitting finger behaviour. In addition properties such as porosity and saturation can be readily up-scaled using statistical methods.%, like the python function that was introduced in the second case of the simulations, that provides fast and reasonable values.
To conclude, up-scaling or averaging the geological model can assist in running reservoir models faster and as close as possible to fine geological model.}

@@

In  this paper, viscous instabilities for different test-cases will be investigated. Adaptive and unstructured mesh will be used to capture the collapse of the interface and the development of the fingers in relation to different values of mobility ratios and different maps of permeabilities. Initially cases that verify

@@

>>>>

\title{ \blue{Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}}

@@

\begin{abstract} 
\KCnote{ \blue{......numerical modelling of multiphase flow in porous media has a wide range of applications in science and engineering. Viscous fingering, a fluid instability can be a major concern regarding enhanced oil-recovery methods (EOR). Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics and tackle the difficulties that commercial resevoir simulators may rise when using low order finite volume/differences methods based on structured grids. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
%Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations. We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.
In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion. Both the evolution of visualisation techniques through the extensive use of parallel computing as well as the advances in modelling and simulation of porous media flows may help to address the importance of this phenomenon which can be found in petroleum, chemical and environmental operations.
The aim of this paper is to extend and further apply the formulation (embedded in the Fluidity software framework) to multiphase flows in heterogeneous porous media combined with either uniform or random generated permeability distribution...}}

@@

In  this paper, viscous instabilities for different test-cases will be investigated. Adaptive and unstructured mesh will be used to capture the collapse of the interface and the development of the fingers in relation to different values of mobility ratios and different maps of permeabilities. Initially test-cases that verify the existence of fingers in plain domain will illustrated follow by more complicated computational domains and finger patterns.

@@

\subsection{Model Validation}
